Shortfall in Health Infrastructure vs Government Funding
========================================================

This is a document showing some results we found by doing EDA(Explorartory Data Analysis) on healthcare datasets from data.gov.in. 

The analysis.R script does some data wrangling and create a table with shortfall, government effort and surplus or shortfall between the two.
```{r}
require(ggplot2)
require(dplyr)
require(reshape2)
source("analysis.R")
```

You can also embed plots, for example:

```{r fig.width=10, fig.height=10}
histogram_Data <- melt(analysis_data, id=c("state_and_ut"))
noTotalsData <- filter(histogram_Data, state_and_ut != c("All India Total"))
qplot(x=value, data=noTotalsData, binwidth=100) + 
  facet_wrap(~variable)
```

This shows us that in the sub centre infrastrucure. Two states are the real outliers. The main problematics states are ..

```{r}
analysis_data %.%
  filter(state_and_ut != c("All India Total"), sc_shortfall > 2000)  %.%
  arrange(desc(sc_shortfall)) %.%
  select(state_and_ut, sc_shortfall)
```


When we add the amount of effort government is putting to improve the matters
```{r}
analysis_data %.%
  filter(state_and_ut != c("All India Total"), sc_shortfall > 2000)  %.%
  arrange(desc(sc_shortfall)) %.%
  select(state_and_ut, sc_shortfall, sc_govt_effort)
```


